Supplementary Information 1

### 2025-06-01 MS2_Krill.Rmd
adding word format chgs

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División NOAA, USA]
  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Richard Methot ^[NOAA, Fishery Science División, NorthWeast, Seattle, USA]
output:
  bookdown::word_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem management objectives in the Western Antarctic Peninsula (WAP). In this study, we implement an integrated length-to-age model that not only allows the incorporation of diverse data sources related to krill population dynamics, but also enables the evaluation of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status. We explored four scenarios that accounted (or not) for ecosystem influences on krill population dynamics, yielding results that align with the expected behavior of a stock assessment framework tailored to this type of species. Performance analyses reveal systematic differences across those scenarios, highlighting increased bias in fishing mortality (F) and spawning stock biomass (SSB) in models that exclude ecosystem variables. Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics. Despite ongoing challenges in incorporating ecosystem variables into integrated model like this, we argue that these factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through pre and post analyses. Overlooking these drivers may result in less accurate conclusions, hence limiting the effectiveness of management recommendations. Integrated modeling approaches that draw on extensive biological, ecologicl, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.

or

Understanding the population dynamics of Antarctic krill (Euphausia superba) is essential for achieving ecosystem-based management objectives in the Western Antarctic Peninsula (WAP), particularly under the mandate of CCAMLR. In this study, we implement an integrated length-to-age assessment model designed to assimilate multiple data sources related to krill biology, fishery operations, and ecological interactions. A key feature of our approach is the explicit inclusion of ecosystem drivers—namely, environmental variability and predation pressure—as intrinsic components of the population dynamics. We evaluate four model scenarios, with and without ecosystem variables, and assess their performance in estimating recruitment, fishing mortality (F), total biomass, and spawning stock biomass (SSB). The comparison reveals systematic biases in models that omit ecological information, resulting in overestimated recruitment and biomass levels. In contrast, scenarios incorporating predator and environmental data yield more precautionary and biologically realistic outcomes, suggesting greater alignment with observed ecosystem variability. Despite ongoing challenges related to the availability, resolution, and integration of ecological data, we argue that environmental and trophic drivers should be embedded directly into the state-space structure of assessment models rather than treated as exogenous factors. Failing to do so may obscure key population processes and compromise the reliability of management advice. Our results underscore the potential of ecosystem-informed, integrated modeling frameworks to support robust, adaptive management strategies that align with CCAMLR’s conservation objectives, while also highlighting the continued need for methodological development and interdisciplinary data collection efforts in the Southern Ocean.

*Keywords: Antarctic krill, population dynamics, length-to-age model, stock-recruitment, ecosystem variables, perfomance model, fisheries management, CCAMLR*

@@

Antarctic krill (*Euphausia superba*) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @Kawaguchi2024; @McBride2021]. Currently, the population, and consecuently, the fishery is predominantly concentrated in the Western Antarctic Peninsula (WAP). This area is also a critical feeding ground for predators such as penguins, seals, and whales, and the primary area for krill harvesting over the past four decades [@Gallagher2023a; @Watters2013; @Hill2016]. At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Lima2013; @Turner2005; @Carrasco2021; @Atkinson2019a; @Morley2020; @McBride2021]. These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective in such a complex and changing environment [@Warwick-Evans2022; @Kawaguchi2024].

Given its ecosystem significance and commercial value, effective management of krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries. Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) that follows an ecosystem-based approach has been developing [@Cavanagh2016; @Constable2023; @Constable2011; @Watters2013]. This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and time. It includes three key components: (i) stock assessments to determine precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a risk assessment framework, now referred to as spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Kawaguchi2024; @Warwick-Evans2022].

To provide scientific advice and recommendations on the status and standing stock of krill, CCAMLR has historically relied on dedicated scientific surveys [@Watters2013; @Green2023; @Hill2019; @Dornam2021]. However, changes in survey design—both methodological and spatial—have raised concerns about the reliability of these estimates, and the high costs associated with such studies have further motivated the search for alternative stock estimation methods. More recently, and as part of a new strategy, CCAMLR has explored a variety of model-based (indirect) approaches to assess krill stock status and support fisheries management decisions [@Hill2019]. Some of these stock assessment models have used a single source of data—either fishery-dependent catches [@Wang2021] or survey-based monitoring programs [@Kinzey2011; @Kinzey2019a; @Green2023]. Nowadays, a model-based approach used for generating management recommendations in CCAMLR is the Generalized Yield Model (GRYM) [@Maschette2020], which allows adjustments to the decision rules to predefined captures set by CCAMLR [@Constable2000; @CCAMLR2021a; @CCAMLR2010]. However, these model-based approaches exhibit limitations, including an effective integration of multiple data sources, incorporation of ecosystem-level variables into population dynamics and/or failure to account for the high spatial and temporal variability intrinsic to krill systems [@Meyer2020; @Trathan2021].

Increasing evidence shows that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Flores2012; @Morley2020; @Saba2014; @Perry2020; @Veytia2021; @Walsh2020; @Murphy2007; @Vorrath2020]. Likewise, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Reisinger2022; @Riaz2023; @Baines2022; @Biuw2024; @Moosa2017]. There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions [REF]. Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Marshall2019; @Crone2019; @Richards2016]. Therefore, any stock assessment frameworks related with new strategy of CCAMLR should explicitly incorporate these ecosystem components to more accurately represent krill population processes.

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data consistency is often affected by changes in monitoring programs and sampling methodologies [REF], while some key regions of the Western Antarctic Peninsula (WAP) lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [REF]. Despite these challenges, substantial data exist on krill and their environment. Fishery monitoring programs provide detailed catch and effort data through onboard scientific observers. Ecosystem monitoring, like the "CCAMLR Ecosystem Monitoring Program" (CEMP), track predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations. Spatially explicit approaches include the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011]. For statistical catch-at-age assessments, key frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013] among others. These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and populations dynamics [@Hoyle2020].

This study aims to integrate different sources of available data on Antarctic krill (*Euphausia superba*) —including fishery operations (catches, effort distribution), biological characteristics (growth rates, reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (environmental forcing and predator-prey interactions)—into a unified stock assessment framework. Specifically, we employ an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting  biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration. The primary objective is to evaluate the statistical performance of alternative model configurations representing different hypotheses about ecosystem influences using a suite of diagnostic tools and performance metrics commonly applied in contemporary stock assessment practices. Our results provide empirical evidence for the value of ecosystem-informed in model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in  model fit, predictive skill and structural uncertainty.

This approach seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach [@CCAMLRWGSAM2024; @Hill2019; @Constable2023; @Zhao2023; @WGEMM2024; @Cavanagh2016; @Constable2011]. Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.

@@

The study area covers the Western Antarctic Peninsula (WAP), corresponding to FAO Subarea 48.1, which hosts a high of krill fishing activity. This is also the region where CCAMLR is actively developing a revised management approach at a finer spatial scale. To enhance the spatial resolution of our analysis of krill population dynamics, we adopted the five Management Units (MUs) (strata) proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig. 1, S1). It is worth noting that these strata were originally developed for management purposes, without explicitly considering biological or population-specific criteria [@Zaldua2024a; @Constable2023; @Watters2013]. However, their subdivision enables a finer-scale analysis of krill data. This increased spatial resolution allows for the detection of heterogeneity among various data sources—whether from fisheries or environmental  variables—thus providing a more detailed understanding of krill population dynamics.

@@

To represent spatial heterogeneity in krill dynamics we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* concept [@Punt2019; @Nielsen2021; @Waterhouse2014]. The areas-as-fleets approach aims treats distinct geographical regions as independent fleets, under the assumption that fishing and survey operations can exert different effects across space. In this approach, each strata within the Western Antarctic Peninsula is treated as a distinct fleet and allows a spatial differentiation in key parameters such as selectivity, catchability (*q*) and abundance indices. This is especially relevant for krill, given the spatial variability in both population processes and fishing effort, given the abundant evidence of spatial heterogeneity [@Perry2019; @Perry2020; @Reiss2020; @SantaCruz2022; @Veytia2021]. By modeling each MU as a separate fleet, we allow the assessment to estimate region-specific parameters that reflect this variability. Survey indices and biological compositions (e.g., length-frequency data) are assigned to their corresponding MU based on sampling location based in @Dornam2021. Similarly, fishery-dependent data—such as catch and composition—are stratified by MU according to the reported fishing effort allowing for regional variation in both abundance and observational processes. Given the difference en the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each strata to account for potential variations in size composition due to spatial heterogeneity or environmental conditions. This disaggregated structure allows the model to reflect small-scale spatial dynamics while maintaining analytical simplicity. The areas-as-fleets approach captures regional variability in both population dynamics and observational processes without the added complexity of explicitly modeling movement [@Punt2019]. The model accounts for interannual variability in growth dynamics, maturity, and reproduction, with a single annual spawning pulse occurring between January and February [@Perry2019; @McBride2021]. Although strata are treated as separate fleets, core population processes—such as recruitment and growth—are modeled collectively, representing the krill stock as a single, interconnected population providing a more realistic representation of krill dynamics population within FAO Subarea 48.1.

@@

Like many invertebrates, krill is a hard to age determination, and this presents challenge for modeling population dynamics, because most modern stock assessments that aim to estimate marine population variables rely on age-structured modeling approaches [@Punt2013]. Although recent advancements have been made [@Kilada2017], aging techniques have historically not been available for krill. Therefore, to develop a stock assessment model for this type of species use length data to approximate age dynamics through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length [@Laslett2004a; @Lee2019; @Lee2024; @Rudd2021; @Zhang2022]. Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Lee2024; @Punt2013]. On the other hand, length data, which are cost-effective and readily available [@Canales2021; @Chong2019], provide valuable insights into krill population structure due to their correlation with age [@Thanassekos2014]. For made this transformation, the models often use an Age-Length Keys (ALKs). An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classed and this method assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024]. The ALK matrix is typically defined as:

@@

where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age. Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches. To date, all model-based krill assessments have used length-to-age frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021; @Green2023]. The resulting ALK is shown in S1 Fig 7.

>>>>

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile, Centro IDEAL, Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::pdf_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

\newpage
\tableofcontents
\newpage

@@

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem-based management fisheries in the Western Antarctic Peninsula. In this study, we implemented an integrated length-to-age model that not only incorporates diverse data sources related to krill population dynamics, but also enables the assessment of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status. We explored four scenarios that either accounted for or ignored ecosystem influences on krill population dynamics, yielding results consistent with the expected behavior of a stock assessment framework tailored to this type of species. Preliminary performance analyses reveal systematic differences across scenarios, highlighting increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB) in models that include ecosystem variables. Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics. Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through pre and post analyses. Overlooking these drivers may lead to less accurate conclusions, hence limiting the effectiveness of management recommendations. Integrated modeling approaches that draw on extensive biological, ecological, fishery, and environmental datasets provide a robust foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use. Such approaches are particularly relevant for ecologically complex systems like the krill fishery in the Western Antarctic Peninsula.

*Keywords: Antarctic krill, population dynamics, length-to-age model, stock-recruitment, ecosystem variables, performance model, fisheries management, CCAMLR*

@@

Antarctic krill (*Euphausia superba*, hereafter krill) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @McBride2021; @Kawaguchi2024]. Currently, the population is predominantly concentrated in the Western Antarctic Peninsula (WAP), being the primary area for krill harvesting over the past four decades and also an area that is also a critical feeding ground for predators such as penguins, seals, and whales [@Watters2013; @Hill2016; @Gallagher2023a]. At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Turner2005; @Lima2013; @Atkinson2019a; @Morley2020; @Carrasco2021; @McBride2021]. These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective in such a complex and changing environment [@Warwick-Evans2022; @Kawaguchi2024].

Given the importance of krill to the ecosystem and its commercial value, effectively managing krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries. Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has been working on a revised ecosystem-based approach [@Constable2011; @Watters2013; @Cavanagh2016; @Constable2023]. This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and temporal scales. It comprises three key components: (i) stock assessments to establish precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Warwick-Evans2022; @Kawaguchi2024].

To provide scientific advice and recommendations on the status and standing stock of krill, CCAMLR has historically relied on dedicated scientific surveys [@Watters2013; @Hill2019; @Dornam2021; @Green2023]. However, changes in survey design—both methodological and spatial—have raised concerns about the reliability of these estimates, while the high costs associated with such surveys have further motivated the search for alternative stock estimation methods. More recently, and as part of a new strategy, CCAMLR has explored a variety of model-based (indirect) approaches to assess krill stock status and support fisheries management decisions [@Hill2019]. Some of these stock assessment models have used a single source of data—either fishery-dependent catches [@Wang2021] or survey-based monitoring programs [@Kinzey2011; @Kinzey2019; @Green2023]. Nowadays, the Generalized Yield Model (GRYM) [@Maschette2020] is a model-based approach used by CCAMLR to generate management recommendations, allowing adjustments to the decision rules to predefined catch limits set by CCAMLR [@Constable2000; @CCAMLR2010; @CCAMLR2021a]. However, model-based approaches exhibit limitations, including an effective integration of multiple data sources, incorporation of ecosystem-level variables into population dynamics, and accounting for the high spatial and temporal intrinsic variability of krill productivity [@Meyer2020; @Trathan2021].

Increasing evidence demonstrates that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021]. Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024]. There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@REF]. Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019]. Therefore, stock assessment frameworks aligned with CCAMLR’s new strategy should explicitly incorporate these ecosystem components to more accurately represent krill population processes.

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@REF], while some key regions of the WAP lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@REF]. Despite these challenges, substantial data exist on krill and associated environments. Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers. Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations. Spatially explicit approaches include the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011]. For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others. These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and population dynamics [@Hoyle2020].

This study aims to integrate various data sources on krill (*Euphausia superba*), including fishery operations (e.g. catches and effort distribution), biological characteristics (e.g. growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g. environmental forcing and predator-prey interactions) into a unified stock assessment framework. Specifically, we use an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration. This approach strengthens the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR's ecosystem-based management approach [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024]. The primary objective is to evaluate the statistical performance of alternative model configurations representing different hypotheses about ecosystem influences using a suite of diagnostic tools and performance metrics commonly applied in contemporary stock assessment practices. Our results provide empirical evidence for the value of ecosystem-informed model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in model fit, predictive skill, and structural uncertainty.

@@

The study area covers the WAP, corresponding to FAO Subarea 48.1, one of the main regions for krill fishing activity. This area is also where CCAMLR is actively revising management approach at a finer spatial scale. To enhance the spatial resolution of our analysis of krill population dynamics, we used the five Management Units (MUs) (strata) proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig. 1, S1). It is worth noting that these strata were originally developed for management purposes, without explicitly considering biological, ecological or population-specific criteria [@Watters2013; @Constable2023; @Zaldua2024a]. Nevertheless, their subdivision enables a finer-scale analysis of krill data, increasing spatial resolution and enabling detection of heterogeneity among various data sources—whether from fisheries or environmental variables—thus providing a more detailed understanding of krill population dynamics.

@@

To represent spatial heterogeneity in krill dynamics, we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* methodology [@Waterhouse2014; @Punt2019; @Nielsen2021]. This approach treats distinct geographical regions as independent fleets, under the assumption that fishing and survey operations may have varying effects across space. Within this approach, each MU of the WAP is treated as a distinct fleet, allowing spatial differentiation in key parameters such as selectivity, catchability (*q*) and abundance indices. This is especially relevant for krill due to documented spatial variability and heterogeneity in both population processes and fishing effort [@Perry2019; @Perry2020; @Reiss2020; @Veytia2021; @SantaCruz2022]. By modeling each MU as a separate fleet, the assessment can estimate region-specific parameters that reflect this variability. Survey indices and biological compositions (e.g., length-frequency data) are assigned to their corresponding MU based on sampling location as defined by [@Dornam2021]. Similarly, fishery-dependent data—such as catch and size composition—are stratified by MU according to the reported fishing effort, allowing for regional variation in both abundance and observational processes. Given known difference in the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each MU to account for potential variations in size composition due to spatial heterogeneity or environmental conditions. This disaggregated structure allows the model to reflect small-scale spatial dynamics while maintaining analytical simplicity. The areas-as-fleets approach effectively incorporates regional variability in both population dynamics and observational processes without the added complexity of explicitly modeling movement [@Punt2019]. The model accounts for interannual variability in growth, maturity, and reproduction, assuming a single annual spawning pulse occurring between January and February [@Perry2019; @McBride2021]. Although MUs are treated as separate fleets, core population processes—such as recruitment and growth—are modeled collectively, representing the krill stock as a single interconnected population. This provides a more realistic representation of the krill population dynamics within Subarea 48.1.

@@

Like many invertebrates, krill present challenges for age determination, which complicates modeling population dynamics. Most modern stock assessments rely on age-structured modeling approaches to estimate marine population parameters [@Punt2013]. Although recent advancements have improved krill aging techniques [@Kilada2017], such methods have historically been unavailable for krill. Consequently, stock assessment for krill typically uses length data to approximate age structure through transition matrices. These matrices estimate the probability that individuals of a given length belong to specific age classes [@Laslett2004; @Lee2019; @Rudd2021; @Zhang2022; @Lee2024]. Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches. To date, all model-based krill assessments have used length-to-age frameworks [@Kinzey2015; @Kinzey2019; @Wang2021; @Green2023]. Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Punt2013; @Lee2024]. On the other hand, length data, which are cost-effective and readily available [@Chong2019; @Canales2021], provide valuable insights into krill population structure due to their correlation with age [@Thanassekos2014]. To perform this transformation, models often use Age-Length Keys (ALKs). An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes. This method assumes both aged and measured individuals are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024]. The ALK matrix is typically defined as follows:

@@

where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age. The resulting ALK is shown in S1 Fig 7.

In the stock assessment catch-at-length model, growth is modeled using a von Bertalanffy growth function lenght-at-age parameterization with parameters L~inf~, k, and coefficient of variation (CV), as outlined in @Mardones2023. This growth relationship is integrated into the model alongside weight-at-length, which is assumed known and calculated using the parameters from @Maschette2020 for krill. The model assumed linear growth below a predefined settlement age of 1.5 years. The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary. Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints. Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age. Maturity was modeled as an age-based logistic function, with the first age of maturity set at 2-3 years. Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$. The model used a direct parameterization without additional offsets between sexes. Regarding selectivity functions, it was modeled using a logistic formulation for the fishery fleet. The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95%). The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width. The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata or "fleet". The survey selectivity was also modeled using a length-based logistic function. Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm. The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values). These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (S1 Fig 6). The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm. assuming that the predator can capture a wider range of krill sizes. These parameters were estimated in later phases, allowing flexibility in their estimation. Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step. The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt model curve like @Kinzey2019, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies. This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@Dong2025; @Kinzey2019; @Perry2020]. The unfished recruitment ($R_0$) was estimated with a mean recruitment of 23 in log-scale, with a prior of 10. The parameter was freely estimated in phase 1. The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability. The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts. Recruitment deviations were estimated to account for interannual variability in recruitment dynamics. The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term. The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years. Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed. Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment. To prevent overparameterization, recruitment deviations were constrained between -5 and 5. Cyclic patterns in recruitment were assumed each tree and five years [@Perry2019; @McBride2021; @Kinzey2019; @Thanassekos2014], and advanced recruitment options were enabled for model flexibility. Life history parameters—such as growth, weight-length relationship, natural mortality, and maturity—were used as priors to model the initial conditions of krill population dynamics, based on data from @Smith2023a, @Maschette2020, and @Kinzey2011. These parameters, along with initial selectivity assumptions, are summarized in Table \@ref(tab:parainit1).

### 2025-06-01 index.Rmd
adding word format chgs

err_long <- err_long %>%
  mutate(Model = case_when(
    Model == "1" ~ "Ref Model: No Env-Predator",
    Model == "2" ~ "S1.1 w/ Predator data",
    Model == "3" ~ "S1.1 w/ Env data",
    Model == "4" ~ "S1.1 w/ Env and Predator data",
    TRUE ~ Model
  ))

@@

## Recruit deviation

```{r, warning=FALSE}
dev1 <- base.model1.1$recruit[8:35,c(1,7)]%>% 
  mutate(Serie = "s1.1")
dev2 <- base.model1.2$recruit[8:35,c(1,7)] %>% 
  mutate(Serie = "s1.2")
dev3 <- base.model1.3$recruit[8:35,c(1,7)] %>% 
  mutate(Serie = "s1.3")
dev4 <- base.model1.4$recruit[8:35,c(1,7)] %>% 
  mutate(Serie = "s1.4")

@@

colores <- c("s1.1" = "#ca0020", 
             "s1.2" = "#f4a582", 
             "s1.3" = "#bababa", 
             "s1.4" = "#404040")
ggplot(df_all, aes(x = year, y = exploitation_rate, color = model)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = colores) +
  labs(
    title = "Exploitation Rate Over Time",
    x = "Year",
    y = "Exploitation Rate",
    color = "Model"
  ) +
  xlim(1997,2020)+
  theme_minimal()
```

>>>>

err_long <- err_long %>%
  mutate(Model = case_when(
    Model == "1" ~ "s1.1",
    Model == "2" ~ "s1.2",
    Model == "3" ~ "s1.3",
    Model == "4" ~ "s1.4",
    TRUE ~ Model
  ))

@@

```{r eval=FALSE, warning=FALSE}
## Recruit deviation
dev1 <- base.model1.1$recruit[8:35,c(1,7)]%>% 
  mutate(Serie = "s1.1")
dev2 <- base.model1.2$recruit[8:35,c(1,7)] %>% 
  mutate(Serie = "s1.2")
dev3 <- base.model1.3$recruit[8:35,c(1,7)] %>% 
  mutate(Serie = "s1.3")
dev4 <- base.model1.4$recruit[8:35,c(1,7)] %>% 
  mutate(Serie = "s1.4")

@@

colores <- c("s1.1" = "#ca0020", 
             "s1.2" = "#f4a582", 
             "s1.3" = "#bababa", 
             "s1.4" = "#404040")
ggplot(df_all, aes(x = year, y = exploitation_rate, color = model)) +
  geom_line() +
  geom_point(size = 3, shape = 16) +
  scale_color_manual(values = colores) +
  scale_x_continuous(
    breaks = seq(1998, 2020, by = 1)  # etiquetas cada 2 años
  ) +
  labs(
    title = "",
    x = "",
    y = "Exploitation Rate",
    color = "Model"
  ) +
  xlim(1997,2020)+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)  # texto vertical
  )
```
